---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

*This is a partial copy from the [original repository](https://github.com/NEvanDis/WM_fitness), made for the Hackathon.  
We kept all the repository the same, except the script folder. There you find only the script `1a_CatFoodExp2021_analysis_fitness.R` which is our focus in the hackathon.*


[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.8276288.svg)](https://doi.org/10.5281/zenodo.8276288) Version of record scripts

# Winter moth individual fitness and population growth
This folder contains all the scripts needed to reproduce the analysis of experimental and field winter moth data belonging to manuscript _Phenological mismatch affects individual fitness and population growth in the winter moth_, published in Proc Roy Soc B: https://doi.org/10.1098/rspb.2023.0414

**NB: The raw data, including both experimental data and field data, can be found on Dryad https://doi.org/10.5061/dryad.m905qfv5p**

&nbsp;

## Authors
Natalie E. van Dis, ORCID ID: 0000-0002-9934-6751

&nbsp;

## Analysis and visualization of Experimental data
R scripts to reproduce the analysis and visualization (incl. manuscript figures) of the 2021 winter moth caterpillar feeding experiment:

### Script: ```2_scripts/1a_CatFoodExp2021_analysis_fitness.R ```
(1) What are the fitness consequences of day to day timing (a)synchrony with budburst?

### Script: ```2_scripts/1b_CatFoodExp2021_analysis_devtime.R ```
(2) Can food quality affect the timing of life stages?

### Script: ```2_scripts/suppl_pupaweight_proxy.R ```
Supplemental: Is pupation weight a good proxy for fecundity?

See ```_src/env_CatFoodExp2021_analysis.txt``` for used R package versions.

&nbsp;

## Analysis and visualization of long-term Field data
### Script: ```2_scripts/2_prep_FieldData.R ```
R script to get trapping effort descriptives and to prep all the field data for analysis.

### Script: ```2_scripts/3_plot_popnum.R ```
R script to reproduce the manuscript figure that visualizes winter moth population dynamics and population phenological mismatch over time (1993-2021) at four locations in the Netherlands.

### Script: ```2_scripts/4_popdyn_analysis.R ```
R script to reproduce the analysis of winter moth population dynamics: how much variation in population growth can be explained by timing mismatch?

See ```_src/env_PopDyn_analysis.txt``` for used R package versions.

## Instructions

The project uses [`renv`](https://rstudio.github.io/renv/index.html) for package management and was run under `r sessionInfo()$running`, [RStudio](https://posit.co/download/rstudio-desktop/) version `r rstudioapi::versionInfo()$version` and `r R.version.string`. To ensure full reproducibility we suggest running the analyses under the same R version using the R Installation Manager [`rig`](https://github.com/r-lib/rig).

To reproduce the project:

1. Clone this repository. For those not familiar with GitHub, click on the green button $\colorbox{darkgreen}{{\color{white}{<> Code}}}$ on the [project main page](https://github.com/SORTEE/CodeStandard) on GitHub and then on `Download ZIP` to download the entire repository, thus unzip it
2. Install RStudio and rig (see instructions [here](https://posit.co/download/rstudio-desktop/) for RStudio and [here](https://github.com/r-lib/rig?tab=readme-ov-file#%EF%B8%8F-installing-rig-) for `rig`)
3. Open a terminal and use the rig commang `rig add` to install `r R.version.string`:

```
rig add 4.1.2
```

4. Open RStudio with the newly install R version using the command

```
rig rstudio 4.1.2
```
5. Go to `File >> Open Project...` and navigate to the project directory to open the R project (`.Rproj` file). Once the R project is is launched, the appropriate version of `renv` should be automatically downloaded and installed into the project library.
6. Call `renv::restore()` to restore the project library locally on the machine. This step will take some time. Note that package installation may require a compiler and a few other command line tools. Make sure they are installed properly (see instructions [here](https://r-pkgs.org/setup.html#setup-tools))
7. Open the script `1a_CatFoodExp2021_analysis_fitness.R` and run it interactively or run the analysis as a whole by calling

```{r, eval = FALSE}
source("2_scripts/1a_CatFoodExp2021_analysis_fitness.R")
```
